Optimal Approximation of Discrete Random Variables \\ for Estimation of Probabilities of Missing Deadlines

### 2018-09-05 Kolmogorov_Approx.tex
n/2 -> n
wrong e_{i,j}
After talking with Liat
no lemma corrected dual problem

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An algorithm for optimal approximation}\label{sec:algEfficient}
In the scope of this section, let $X$ be a given random variable with a finite support of size $n$, and let  $0<m\leq n$ be a given complexity bound. The section evolves by presenting two algorithms for finding an optimal $m$-approximation of $X$. The first algorithm is $binsearchApprox(X,m)$ (Algorithm~\ref{alg:naive}) which runs in time of $O(n^2log(n))$ and the second one $linApprox(X,m)$ (Algorithm~\ref{alg:linear}) which runs in time of $O(nlog(n))$. We present both $binsearchApprox(X,m)$ and $linApprox(X,m)$ since Algorithm~\ref{alg:linear} is based on Algorithm~\ref{alg:naive} but carry major run-time improvement.

We begin our dissection by presenting a useful fact that it is enough to limit our search to approximations $X'$s such that $\support(X') \subseteq \support(X)$:

\begin{lemma}\label{lem:supContained}
	For every random variable $X''$ there is a random variable $X'$ such that $\support(X') \subseteq \support(X)$ and $d_{K}(X,X')\leq d_{K}(X,X'')$.
\end{lemma}

\begin{proof}
	Let $\{x_1 <\cdots<x_n\} = \support(X)$, and let $x_0 = -\infty, x_{n+1}=\infty$. Consider the random variable $X'$ whose probability mass function is
	$f_{X'}(x_i) = P(x_{i-1} < X'' \leq x_i)$ for $i=1,\dots,n-1$,  $f_{X'}(x_n) = P(x_{n-1} < X'')$, and $F_{X'}(x)=0$ if $x\notin \support(X)$.  Since we only "shifted" the probability mass of $X''$ to the support of $X$, we have that $f_{X'}$ is a probability mass function and therefore $X'$ is well defined. 	
	By construction, $|F_{X}(x_i)-F_{X'}(x_i)| = |F_{X}(x_i)-F_{X''}(x_i)|$ for every  $0 <  i < n$. For $i=n$ we have $|F_{X}(x_n)-F_{X'}(x_n)| = |1-1|=0$.
	Since $|F_{X}(x)-F_{X'}(x)| = |F_{X}(x_i)-F_{X'}(x_i)|$ for every  $0\leq   i  \leq  n$  and $x_i<x<x_{i+1}$, we have that $d_K(X,X')=max_{i}|F_{X}(x_i)-F_{X'}(x_i)|\leq max_{i}|F_{X}(x_i)-F_{X''}(x_i)|\leq d_K(X,X'')$.
\end{proof}

Both of the algorithms presented in this section are using the dual problem in order to find the optimal $m$-approximation of $X$. The dual problem is: given a random variable $X$ with $|X| = n$, and some $0 \leq \varepsilon \leq 1$, find a new random variable $X'$ such that $|X'|\leq m$ where $0< m \leq n$ and $d_K(X,X')$ is minimal.

@@

\begin{algorithm}
	\DontPrintSemicolon
	Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$.\;
    $f \gets 0$, $l \gets n$+1\; 
    \While{$c_{f +1} \leq \varepsilon$}{$f \gets f + 1$}

@@

	\caption{$dual(X,\varepsilon)$}   
	\label{alg:dual}
\end{algorithm}

@@

		\lIf {$m'  > m$}
		{
			$j \gets j - 1$
		}
		\lIf {$m' \leq m$} {
			$i \gets i + 1$, $S \gets S \cup (m',e)$			
		}
	}
	$e \gets \min\{e\colon (m',e) \in \underset{(m',e)\in S}{\operatorname{arg\,max}}\, m'\}$\;
	\Return{$dual(X,e)$}

@@

		$S \gets  \{ ((1, 1)),(w,h)) \}$, $S' \gets \emptyset$\;

@@

	Let 
	$$e_{i,j} = \begin{cases}
	c_j   & \text{if } i > \lfloor n/2 \rfloor; \\
	1-c_i & \text{if } j \geq \lceil n/2 \rceil; \\
	(c_{n+1-j} {-} c_i)/2  & \text{otherwise}.
	\end{cases}$$

@@

We describe below several experiments that show how $\KlmApprox$ performs in practice in different applications and domains.
All algorithms were implemented in Python and the experiments were executed on a hardware comprised of an Intel i5-6500 CPU @ 3.20GHz processor and 8GB memory. The algorithms of Cohen et al. were taken "as is" from in the supplementary material to~\cite{cohen2015estimating}.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{An algorithm for optimal approximation}\label{sec:algEfficient}
In the scope of this section, let $X$ be a random variable such that $|X|=n$, and let  $0<m\leq n$. We present algorithms for finding optimal $m$-approximations of $X$. The first algorithm is $binsearchApprox(X,m)$ (Algorithm~\ref{alg:naive}) which runs in time of $O(n^2\log(n))$. The second algorithm, $linApprox(X,m)$ (Algorithm~\ref{alg:linear}), runs in time of $O(n\log(n))$. 
%We present both $binsearchApprox(X,m)$ and $linApprox(X,m)$ since Algorithm~\ref{alg:linear} is based on Algorithm~\ref{alg:naive} but carry major run-time improvement.

%We begin our dissection by presenting a useful fact that it is enough to limit our search to approximations $X'$s such that $\support(X') \subseteq \support(X)$:
%
%\begin{lemma}\label{lem:supContained}
%	For every random variable $X''$ there is a random variable $X'$ such that $\support(X') \subseteq \support(X)$ and $d_{K}(X,X')\leq d_{K}(X,X'')$.
%\end{lemma}
%
%\begin{proof}
%	Let $\{x_1 <\cdots<x_n\} = \support(X)$, and let $x_0 = -\infty, x_{n+1}=\infty$. Consider the random variable $X'$ whose probability mass function is
%	$f_{X'}(x_i) = P(x_{i-1} < X'' \leq x_i)$ for $i=1,\dots,n-1$,  $f_{X'}(x_n) = P(x_{n-1} < X'')$, and $F_{X'}(x)=0$ if $x\notin \support(X)$.  Since we only "shifted" the probability mass of $X''$ to the support of $X$, we have that $f_{X'}$ is a probability mass function and therefore $X'$ is well defined. 	
%	By construction, $|F_{X}(x_i)-F_{X'}(x_i)| = |F_{X}(x_i)-F_{X''}(x_i)|$ for every  $0 <  i < n$. For $i=n$ we have $|F_{X}(x_n)-F_{X'}(x_n)| = |1-1|=0$.
%	Since $|F_{X}(x)-F_{X'}(x)| = |F_{X}(x_i)-F_{X'}(x_i)|$ for every  $0\leq   i  \leq  n$  and $x_i<x<x_{i+1}$, we have that $d_K(X,X')=max_{i}|F_{X}(x_i)-F_{X'}(x_i)|\leq max_{i}|F_{X}(x_i)-F_{X''}(x_i)|\leq d_K(X,X'')$.
%\end{proof}

The algorithms presented in this section are using a dual problem in order to find the optimal $m$-approximation of $X$. The dual problem is: given a random variable $X$ and $0 \leq \varepsilon \leq 1$, find a new random variable $X'$ such that $d_k(X,X')\leq \varepsilon$ and $|X'|$ is minimal.

@@

\begin{algorithm}
	\DontPrintSemicolon
	%Let $\{(x_i, c_i)\}_{i=1}^n$ be such that $c_i = Pr(X \leq x_i)$ and $\supp(X)=\{x_1 <\cdots < x_n\}$.\;
    $f \gets 0$, $l \gets n$+1\; 
    \While{$c_{f +1} \leq \varepsilon$}{$f \gets f + 1$}

@@

	\caption{$dual(\{(x_i, c_i)\}_{i=1}^n,\varepsilon)$  }   
	\label{alg:dual}
\end{algorithm}

@@

		\lIf {$m'  \leq m$}
		{
			$j \gets j - 1$
		}
		\lIf {$m' > m$} {
			$i \gets i + 1$, $S \gets S \cup (m',e)$			
		}
	}
	$e \gets \min\{e\colon (m',e) \in \underset{(m',e)\in S}{\operatorname{arg\,max}}\, m'\}$\;
	\Return{$dual(X,e)$}

@@

		$S \gets  \{ ((1, 1),(w,h)) \}$, $S' \gets \emptyset$\;

@@

	Let 
	$$e_{i,j} = \begin{cases}
	c_j   & \text{if } j \leq n \wedge  i > n; \\
	1-c_i & \text{if } i \leq n \wedge  j \geq n; \\
	%(c_{n+1-j} {-} c_i)/2  & \text{if } i \leq n \wedge i{+}j\leq n{+}1; \\

	%(c_{n+1-i} {-} c_i)/2  & \text{if } j \leq n \wedge i{+}j\leq n; \\

	(c_{p-1 +i} {-} c_{p+1-j})/2  & \text{if } i \leq n \wedge i{+}j\leq n{+}1; \\

	(c_{p-1 +i} {-} c_{p+1-i})/2  & \text{if } i \leq n \wedge i{+}j\leq n{+}1; \\

	\infty  & \text{otherwise}.
	\end{cases}$$

@@

We describe below several experiments that show how $\KlmApprox$ performs in practice in different applications and domains.
All algorithms were implemented in Python and the experiments were executed on a hardware comprised of an Intel i5-6500 CPU @ 3.20GHz processor and 8GB memory. The algorithms of Cohen et al. were taken ``as is" from in the supplementary material to~\cite{cohen2015estimating}.
